Robust Quantum Optimal Control with Trajectory Optimization

### 2021-03-22 paper/3fluxonium.tex
jens all edits plus some thomas revisions

\section{QOC for the Fluxonium \label{sec:fluxonium}}
In the following, we optimize quantum gates
(unitary transformations) for the superconducting fluxonium qubit--a promising
building block for quantum computers due to its high
coherence times
\cite{earnest2018realization, lin2018demonstration,
  manucharyan2009fluxonium, nguyen2019high,
  zhang2020universal}.
In this section, we use the trajectory optimization
formalism \eqref{eq:gcostfun}-\eqref{eq:eqcon}
to define the optimization problem \eqref{eq:costfun}-\eqref{eq:ic_con},
which we extend in subsequent sections to account
for experimental error channels.
To high accuracy, we approximate the Hamiltonian near the flux-frustration
point as a two-level system:
\begin{align}
  H/h &= f_{q} \frac{\sigma_{z}}{2} + a(t) \frac{\sigma_{x}}{2},
  \label{eq:hamiltonian}
\end{align}
where $f_{q}$ is the qubit frequency at the flux-frustration point,
$a(t)$ is the flux offset from the flux-frustration point,
$h$ is Planck's constant, and $\sigma_{z}, \sigma_{x}$
are Pauli matrices. Although the coherent dynamics can be modeled as a two-level
system, our noise model, experimental constraints, and system parameters
consider the full system, and they are representative of the fluxonium
presented in \cite{zhang2020universal}.

First, we outline the constraints for the fluxonium gate problem.
All gates presented in this work satisfy these constraints to
a maximum violation of $\sim 10^{-8}$.
Casting this problem as a multi-state transfer, the initial conditions on
the states are $\ket*{\psi^{0}_{1}} = \ket*{0}$, $\ket*{\psi^{1}_{1}} = \ket*{1}$
\eqref{eq:istate_con}
where the superscript is an index $i \in \{0, 1\}$,
and the subscript indicates the time step $k = 1$.
The states at the final time step are constrained to be
the target states $\ket*{\psi^{i}_{N}} = \ket*{\psi^{i}_{f}} \equiv
U \ket*{\psi^{i}_{1}} \ \forall \ i$
\eqref{eq:tstate_con} where $U = X/2, Y/2, Z/2$ is the target gate.
Furthermore, we impose the normalization constraint
${\lvert \braket*{\psi^{i}_{k}}{\psi^{i}_{k}} \rvert}^{2} = 1 \ \forall \ i,k$
\eqref{eq:statenorm_con}
to ensure the solver does not take advantage of discretization errors in numerical integration.
To refer to the discrete moments of the flux, we introduce the notation
$\int^{t_{k}}_{t_{1}} a(t) \ \mathrm{d}t \equiv \int_{t} a_{k}$,
$a(t_{k}) \equiv a_{k}$,
$\pdv*[n]{a(t)}{t} \lvert_{t = t_{k}} \equiv \partial^{n}_{t} a_{k}$.
We impose the zero net flux constraint $\int_{t} a_{N} = 0$
\eqref{eq:znf_con}
which mitigates the inductive drift ubiquitous in flux-bias lines
\cite{battistel2019fast, krantz2019quantum, zhang2020universal}.
The flux is constrained by $\lvert a_{k} \rvert \leq 0.5 \ \textrm{GHz} \ \forall \ k$
\eqref{eq:amp_con}.
Above $0.5$ GHz, we observe poor experimental agreement with the two-level approximation \eqref{eq:hamiltonian}.
We also enforce the boundary condition $a_{1} = a_{N} = 0$ \eqref{eq:bound_con}
so the gates may be concatenated arbitrarily. Additionally,
we have the initial condition $\int_{t} a_{1} = \partial_{t} a_{1} = 0$
\eqref{eq:ic_con}.

The augmented control and augmented state are:
\begin{equation}
  \begingroup
  \renewcommand*{\arraystretch}{1.3}
  u_{k} = \begin{bmatrix} \partial^{2}_{t} a_{k} \end{bmatrix}, \quad
  x_{k} = \begin{bmatrix} \ket{\psi^{0}_{k}} \\ \ket{\psi^{1}_{k}}
    \\ \int_{t} a_{k} \\ a_{k} \\ \partial_{t} a_{k} \end{bmatrix}.
  \endgroup
  \label{eq:astatecontrols}
\end{equation}
The time evolution of these variables are encoded
in coupled, first-order, differential equations
which are integrated in the discrete dynamics function \eqref{eq:dyn_con}.
We integrate the states according to the TDSE \eqref{eq:tdse} and the
fluxonium Hamiltonian \eqref{eq:hamiltonian}.
The ALTRO implementation we use does not currently
support complex numbers, so we repesent the states
in the isomorphism $\mathcal{H}(\mathbb{C}^{n})
\cong \mathcal{H}(\mathbb{R}^{2n})$ given in \cite{leung2017speedup},
\begin{equation}
  H \ket{\psi} \cong \begin{bmatrix} H_{\textrm{re}} & -H_{\textrm{im}}
    \\ H_{\textrm{im}} & H_{\textrm{re}}\end{bmatrix}
  \begin{bmatrix} \ket{\psi}_{\textrm{re}} \\ \ket{\psi}_{\textrm{im}}\end{bmatrix}.
  \label{eq:isomorphism}
\end{equation}

The cost function at each time step is
$\ell_{k}(x_{k}, u_{k}) = (x_{k} - x_{f})^{T} Q_{k} (x_{k} - x_{f}) + u^{T}_{k} R_{k} u_{k}$
where $Q_{k}$ and $R_{k}$ are positive-definite diagonal matrices we supply.
The $Q_{k}$ term
penalizes deviations from the target augmented state $x_{f}$,
which is given by the constraints we have imposed on
$\ket*{\psi^{i}_{N}}$, $\int_{t} a_{N}$, and $a_{N}$ in addition to
$\partial_{t} a_{f} = 0$.
The $R_{k}$ term penalizes the norm of $\partial^{2}_{t} a_{k}$,
smoothing the flux to mitigate high-frequency AWG transitions.
Stated succinctly, the optimization problem takes the form:
\begin{mini!}[2] 
  {x_{1:N}, u_{1:N\text{-}1}}{\sum_{k=1}^N {(x_k-x_f)}^{T} Q_k (x_k-x_{f})
    + \sum_{k=1}^{N-1} {u_k}^{T} R_k u_{k}}{}{} \label{eq:costfun}
  \addConstraint{x_{k+1}}{= f(x_k, u_k) \ \forall \ k}  \label{eq:dyn_con}
  \addConstraint{\ket{\psi^{0}_{1}} = \ket{0}, \ket{\psi^{1}_{1}} = \ket{1}} \label{eq:istate_con}
  \addConstraint{\ket{\psi^{i}_{N}} = \ket{\psi^{i}_{f}}
    \ \forall \  i} \label{eq:tstate_con}
  \addConstraint{{\lvert \braket{\psi^{i}_{k}}{\psi^{i}_{k}}
      \rvert}^{2} = 1 \ \forall \ i, k} \label{eq:statenorm_con}
  \addConstraint{{\textstyle \int_{t}} a_{N} = 0} \label{eq:znf_con}
  \addConstraint{|a_{k}| \leq 0.5 \ \textrm{GHz} \ \forall \ k} \label{eq:amp_con}
  \addConstraint{a_{1} = a_{N} = 0} \label{eq:bound_con}
  \addConstraint{{\textstyle \int_{t}} a_{1} = \partial_{t} a_{1} = 0}. \label{eq:ic_con}
\end{mini!}

Next, we remark on our problem formulation.
We penalize the squared error of the final state and the target
state \eqref{eq:costfun}, rather than their infidelity,
because the Hessian of the squared error cost function is diagonal--which
makes matrix multiplications fast--and we wish to optimize $Z/2$ gates,
which requires a metric that is sensitive to global phases for the initial
states $\ket{0}$ and $\ket{1}$.
We penalize the squared error at all time steps
because it benefits the iLQR solving stage \cite{Jackson2020altroc},
not to incentivize early achievement
of the desired gate \cite{leung2017speedup}.

In addition to the target state cost function, we
impose the target state constraint \eqref{eq:tstate_con}, which
says that the final state must match the target state, including its global phase,
up to our chosen maximum constraint violation $\sim 10^{-8}$.
If we did not impose this constraint, the optimizer would be
allowed to sacrifice the closed-system gate error to achieve better
performance on the other cost functions, which is undesirable.
To enforce a constraint in standard QOC frameworks,
the prefactor for the constraint function is manually increased
between separate optimization instances until the constraint is satisfied
\cite{heeres2017implementing, leung2017speedup, reinhold2019controlling},
which becomes infeasible for more than one constraint.
ALM automates these prefactor updates to find
a solution that satisfies all of the given constraints.
Hence, ALTRO's ability to handle multiple constraints makes it
an attractive solver for QOC problems.

In extraordinarily difficult cases of
QOC \cite{abdelhafez2020universal}, it may be impossible
to obey the physics of the system and achieve the desired gate;
equivalently, the dynamics constraint \eqref{eq:dyn_con}
and the target state constraint \eqref{eq:tstate_con} may be mutually unsatisfiable.
In this case, the prefactors for the constraint function terms
in the ALM objective will tend to infinity--leading to numerical instability--and the
optimization will not converge. To maintain a constrained approach in this situation,
the maximum constraint violation for the target state constraint can be raised
to a level commensurate with the minimum acceptable gate error.

>>>>

\section{QOC for the Fluxonium \label{sec:fluxonium}}
In the following, we optimize quantum gates
(unitary transformations) \textcolor{blue}{[for a physics journal, you will want as much handholding and explanation for the robotics community language as possible; bu contrast, clarifying that a quantum gate is a unitary tranformation seems out of place here.]} for the superconducting fluxonium qubit -- a promising
building block for quantum computers due to its high
coherence times
\cite{earnest2018realization, lin2018demonstration,
  manucharyan2009fluxonium, nguyen2019high,
  zhang2020universal}.
In this section, we use the trajectory optimization
formalism \eqref{eq:gcostfun}-\eqref{eq:eqcon}
to define the optimization problem \eqref{eq:costfun}-\eqref{eq:ic_con},
which we extend in subsequent sections to account
for experimental error channels.
To high accuracy, we can approximate the fluxonium Hamiltonian near the flux-frustration
point by a two-level system:
\begin{align}
  H/h &= f_{q} \frac{\sigma_{z}}{2} + a(t) \frac{\sigma_{x}}{2}.
  \label{eq:hamiltonian}
\end{align}
Here, $f_{q}$ is the qubit frequency at the flux-frustration point,
$a(t)$ is the control governing the flux offset from the flux-frustration point,
$h$ is Planck's constant, and $\sigma_{z}, \sigma_{x}$
are Pauli matrices. Although the coherent dynamics can be described based on this two-level
system model, our noise model, experimental constraints, and system parameters
consider the full system, and they are representative of the fluxonium
presented in \cite{zhang2020universal}.

First, we outline the constraints for the fluxonium gate problem.
Casting this problem in terms of a multi-state transfer problem, we fix as the initial states
 $\ket*{\psi^{0}_{1}} = \ket*{0}$, $\ket*{\psi^{1}_{1}} = \ket*{1}$
\eqref{eq:istate_con}
where the superscript $i \in \{0, 1\}$ is the state index,
and the subscript labels the time step -- here, $k = 1$.
The states at the final time step are constrained to be
the target states $\ket*{\psi^{i}_{N}} = \ket*{\psi^{i}_{f}} \equiv
U \ket*{\psi^{i}_{1}} \ \forall \ i$
\eqref{eq:tstate_con} with $U = X/2, Y/2, Z/2$ denoting the target gate.
Furthermore, we impose the normalization constraint
${\lvert \braket*{\psi^{i}_{k}}{\psi^{i}_{k}} \rvert}^{2} = 1 \ \forall \ i,k$
\eqref{eq:statenorm_con}
to ensure the solver does not take advantage of discretization errors in numerical integration.
To refer to the discrete moments of the flux, we introduce the notation
$\int^{t_{k}}_{t_{1}} a(t) \ \mathrm{d}t \equiv \int_{t} a_{k}$,
$a(t_{k}) \equiv a_{k}$,
$\pdv*[n]{a(t)}{t} \lvert_{t = t_{k}} \equiv \partial^{n}_{t} a_{k}$. \textcolor{blue}{[Why is this a partial rather than a single-variable derivative?]}
We impose the zero net-flux constraint $\int_{t} a_{N} = 0$
\eqref{eq:znf_con}
which mitigates the inductive drift ubiquitous in flux-bias lines
\cite{battistel2019fast, krantz2019quantum, zhang2020universal}.
The flux is constrained by $\lvert a_{k} \rvert \leq 0.5 \ \textrm{GHz} \ \forall \ k$
\eqref{eq:amp_con}.
Above $0.5$ GHz, we observe poor experimental agreement with the two-level approximation \eqref{eq:hamiltonian}.
We also enforce the boundary condition $a_{1} = a_{N} = 0$ \eqref{eq:bound_con} \textcolor{blue}{[Thomas, this would actually be consistent with my suggestion to set $\mathbf{u}_N$ to zero?]}
so the gates may be concatenated arbitrarily. Additionally,
we have the initial condition $\int_{t} a_{1} = \partial_{t} a_{1} = 0$
\eqref{eq:ic_con}. All gates presented in this work satisfy these constraints to
a maximum violation of $\sim 10^{-8}$.

The augmented control and augmented state are:
\begin{equation}
  \begingroup
  \renewcommand*{\arraystretch}{1.3}
  \mathbf{u}_{k} = \partial^{2}_{t} {a}_{k} , \quad
  \mathbf{x}_{k} = \begin{bmatrix} \ket{\psi^{0}_{k}} \\ \ket{\psi^{1}_{k}}
    \\ \int_{t} a_{k} \\ a_{k} \\ \partial_{t} a_{k} \end{bmatrix}.
  \endgroup
  \label{eq:astatecontrols}
\end{equation}
(We maintain vector notation for $\mathbf{u}_k$ in anticipation of augmentation with additional components in Section \ref{sec:longitude}.)
The time evolution of these variables is governed by 
 coupled first-order differential equations
which are encoded in the discrete dynamics function \eqref{eq:dyn_con}.
We integrate the states according to the TDSE \eqref{eq:tdse} based on the
fluxonium Hamiltonian \eqref{eq:hamiltonian}.
Since the ALTRO implementation we use does not currently
support complex numbers, we follow \cite{leung2017speedup} in  representing the states
via the isomorphism $\mathcal{H}(\mathbb{C}^{n})
\cong \mathcal{H}(\mathbb{R}^{2n})$ given in \cite{leung2017speedup},
\begin{equation}
  H \ket{\psi} \cong \begin{bmatrix} H_{\textrm{re}} & -H_{\textrm{im}}
    \\ H_{\textrm{im}} & H_{\textrm{re}}\end{bmatrix}
  \begin{bmatrix} \ket{\psi}_{\textrm{re}} \\ \ket{\psi}_{\textrm{im}}\end{bmatrix}.
  \label{eq:isomorphism}
\end{equation}

The cost function at each time step is
$\ell_{k}(\mathbf{x}_{k}, \mathbf{u}_{k}) = (\mathbf{x}_{k} - \mathbf{x}_{f})^{T} Q_{k} (\mathbf{x}_{k} - \mathbf{x}_{f}) + \mathbf{u}^{T}_{k} R_{k} \mathbf{u}_{k}$
where $Q_{k}$ and $R_{k}$ are positive-definite diagonal matrices we supply. \textcolor{blue}{[``We supply" -- sure, but not very helpful. How are they defined/constructed specifically?]}
The $Q_{k}$ term
penalizes deviations from the target augmented state $\mathbf{x}_{f}$,
which is given by the constraints we have imposed on
$\ket*{\psi^{i}_{N}}$, $\int_{t} a_{N}$, and $a_{N}$ in addition to
$\partial_{t} a_{f} = 0$. \textcolor{blue}{[Should this have been $a_N$ rather than $a_f$?]}
The $R_{k}$ term penalizes the norm of $\partial^{2}_{t} a_{k}$,
smoothing the flux to mitigate high-frequency AWG transitions.
Stated succinctly, the optimization problem takes the form:
\begin{mini!}[2] 
  {\substack{\mathbf{x}_1,\ldots,\mathbf{x}_N \\ \mathbf{u}_1,\ldots,\mathbf{u}_{N-1}}}{\sum_{k=1}^N {(\mathbf{x}_k-\mathbf{x}_f)}^{T} Q_k (\mathbf{x}_k-\mathbf{x}_{f})
    + \sum_{k=1}^{N-1} {\mathbf{u}_k}^{T} R_k \mathbf{u}_{k}}{}{} \label{eq:costfun}
  \addConstraint{\mathbf{x}_{k+1}}{= \mathbf{f}(\mathbf{x}_k, u_k) \quad \forall \ k}  \label{eq:dyn_con}
  \addConstraint{\ket{\psi^{0}_{1}} = \ket{0}, \ket{\psi^{1}_{1}} = \ket{1}} \label{eq:istate_con}
  \addConstraint{\ket{\psi^{i}_{N}} = \ket{\psi^{i}_{f}}
    \quad \forall \  i} \label{eq:tstate_con}
  \addConstraint{{\lvert \braket{\psi^{i}_{k}}{\psi^{i}_{k}}
      \rvert}^{2} = 1 \quad \forall \ i, k} \label{eq:statenorm_con}
  \addConstraint{{\textstyle \int_{t}} a_{N} = 0} \label{eq:znf_con}
  \addConstraint{|a_{k}| \leq 0.5 \ \textrm{GHz} \quad \forall \ k} \label{eq:amp_con}
  \addConstraint{a_{1} = a_{N} = 0} \label{eq:bound_con}
  \addConstraint{{\textstyle \int_{t}} a_{1} = \partial_{t} a_{1} = 0}. \label{eq:ic_con}
\end{mini!}

Next, we remark on our problem formulation.
We penalize the squared error of the final state and the target
state \eqref{eq:costfun} \textcolor{blue}{[referencing (6a) is puzzling here unless you specify $R_k$; how is ``squared error" different from infidelity?]}, rather than their infidelity,
because the Hessian of the squared-error cost function is diagonal -- which
makes matrix multiplications fast -- and we wish to optimize $Z/2$ gates,
which requires a metric that is sensitive to global phases for the initial
states $\ket{0}$ and $\ket{1}$.
We penalize the squared error at all time steps
because it benefits the iLQR solving stage \cite{Jackson2020altroc},
not to incentivize early achievement \textcolor{blue}{[NOT to incentivize or to incentivize??]}
of the desired gate \cite{leung2017speedup}.

In addition to the target-state cost function, we
impose a target-state constraint \eqref{eq:tstate_con} which  requires the final state to match the target state, including its global phase,
up to our chosen maximum constraint violation $\sim 10^{-8}$.
If we did not impose this constraint, the optimizer would be
allowed to sacrifice the closed-system gate error to achieve better
performance on the other cost functions, which is undesirable.
To enforce a constraint in standard QOC frameworks,
the prefactor for the constraint function is manually increased
between separate optimization instances until the constraint is satisfied
\cite{heeres2017implementing, leung2017speedup, reinhold2019controlling},
which becomes infeasible for more than one constraint.
ALM automates these prefactor updates to find
a solution that satisfies all of the given constraints.
Hence, ALTRO's ability to handle multiple constraints makes it
an attractive solver for QOC problems.

In extraordinarily difficult cases of
QOC \cite{abdelhafez2020universal}, it may be impossible
to obey the physics of the system and achieve the desired gate;
equivalently, the dynamics constraint \eqref{eq:dyn_con}
and the target-state constraint \eqref{eq:tstate_con} may be mutually unsatisfiable.
In this case, the prefactors for the constraint function terms
in the ALM objective will tend to infinity -- leading to numerical instability -- and the
optimization will not converge. To maintain a constrained approach in this situation,
the maximum constraint violation for the target-state constraint can be raised
to a level commensurate with the minimum acceptable gate error.

### 2021-03-22 paper/4longitude.tex
jens all edits plus some thomas revisions

The integrated depolarization rate is given by,
\begin{equation}
  D_{1}(t) = \int_{0}^{t} T_{1}^{-1}(a(t^{\prime})) dt^{\prime}.
\end{equation}
For the gates we consider here, where the gate time is small compared to $T_{1}$,
the integrated depolarization rate is proportional to the probability of a depolarization event.
Additionally, the integrated depolarization rate is a reasonable proxy for the gate error incurred
because depolarization errors are incoherent--they increase monotonically in time without interference.
The integrated depolarization rate is appended to the augmented state \eqref{eq:astatecontrols}
and its norm is penalized in the objective \eqref{eq:costfun} by setting
the corresponding element of the target augmented state to zero.
$T_{1}(a_{k})$ is obtained at each time step by evaluating
a spline fit to experimental data of the form $\{(a, T_{1})\}$,
see Figure \ref{fig:longitudeb}.

Additionally, a master equation approach would require adding
density matrices of size $n \times n$
to the augmented state, and a quantum trajectory approach
would require adding many states of size $n$ to the augmented state,
where $n$ is the dimension of the Hilbert space.
The integrated depolarization rate is a single
real number; thus, the computational complexity of this
depolarization model does not scale
with the dimension of the Hilbert space.

We allow the optimizer to tune the gate time by
making the duration between time steps
a decision variable \cite{howell2019altro}. 
The square root of the duration $\sqrt{\Delta t_{k}}$
is appended to the augmented control \eqref{eq:astatecontrols}
and its square $\lvert \Delta t_{k} \rvert$ is used
for integration in the discrete dynamics function \eqref{eq:dyn_con}.
To ensure numerical stability, we constrain
the bounds of the duration.

@@

The flux pulses for the numerical gates
are approximately periodic
with amplitudes $\sim 0.2 \textrm{GHz}$, see Figure \ref{fig:longitudea}.
They are reminiscent of the analytically determined Floquet operations
for a fluxonium described in \cite{huang2020engineering}
and realized in \cite{mundada2020floquet}.
The numerical gate times are greater
than the analytic gate times, but the
numerical flux pulses
spend more time at higher flux values, achieving higher $T_{1}$ times on average,
see Figure \ref{fig:longitudeb}.
The single gate errors for both the analytic and numerical gates are
less than $10^{-4}$, which makes them sufficient for quantum error correction--a
prerequisite for fault-tolerant quantum
computing \cite{aharonov2008fault, knill2005quantum, gottesman1997stabilizer}.
However, the numerical gates achieve single gate errors
$\sim 5$ times less than those for the analytic gates,
which tracks closely with their relative improvement
on the integrated depolarization rate metric, see Appendix \ref{appendix:longitude}.
This single gate error advantage corresponds to a
significant reduction in error correction resources
\cite{paetznick2014resource, suchara2013comparing}.
Furthermore, for successive gate applications,
the gate error due to depolarization is approximately linear
in the gate count, which we expect for $t \ll T_{1}$, see Figure \ref{fig:longitudec}.
The gate error reduction for large gate counts is
important for noisy, intermediate-scale quantum (NISQ)
applications. These improvements are significant for the constraints
we have imposed on the gates,
and do not represent a fundamental limit to the optimization
methods we have employed.

>>>>

The integrated depolarization rate is given by,
\begin{equation}
  D_{1}(t) = \int_{0}^{t} T_{1}^{-1}[a(t^{\prime})] dt^{\prime}.
\end{equation}
For the gates we consider here, where the gate time is small compared to $T_{1}$,
the integrated depolarization rate is proportional to the probability of a depolarization event.
Additionally, the integrated depolarization rate is a reasonable proxy for the gate error incurred
because depolarization errors are incoherent -- they increase monotonically in time without interference.
The integrated depolarization rate is appended to the augmented state \eqref{eq:astatecontrols}
and its norm is penalized in the objective \eqref{eq:costfun} by setting
the corresponding element of the target augmented state to zero.
$T_{1}(a_{k})$ is obtained at each time step by evaluating
a spline fit to experimental data of the form $\{(a, T_{1})\}$ \textcolor{blue}{[What's the meaning of the nested parentheses here? Is the outer supposed to be a set? Then elements would need to be indexed by $k$.]},
see Figure \ref{fig:longitudeb}.
\textcolor{blue}{Phys. Rev. will want this to read 1(b). Not sure how to change this the way you are doing automatic referencing.}

An alternative master equation approach would require adding
density matrices of size $n \times n$
to the augmented state, and a quantum trajectory approach
would require adding many states of size $n$ to the augmented state,
where $n$ is the dimension of the Hilbert space.
By contrast, the integrated depolarization rate is a single
real number; thus, the computational complexity of this
depolarization model does not scale
with the dimension of the Hilbert space.

\textcolor{blue}{[abrupt start of a new paragraph; provide transition/motivation for what's next; the entire following paragraph feels a bit out of place/inserted after the fact.]}
We allow the optimizer to tune the gate time by
making the duration between time steps
a decision variable \cite{howell2019altro}. 
The square root of the duration $\sqrt{\Delta t_{k}}$
is appended to the augmented control \eqref{eq:astatecontrols}
and its square $\lvert \Delta t_{k} \rvert$ is used
for integration in the discrete dynamics function \eqref{eq:dyn_con}.
To ensure numerical stability, we constrain
the bounds of the duration.

@@

The flux pulses for the numerical gates
are approximately periodic
with amplitudes $\sim 0.2 \textrm{GHz}$, see Figure \ref{fig:longitudea}.
They are reminiscent of the analytically determined Floquet operations
for a fluxonium described in \cite{huang2020engineering}
and realized in \cite{mundada2020floquet}.
The numerical gate times are greater
than the analytic gate times, but the
numerical flux pulses
spend more time at higher flux values, achieving higher $T_{1}$ times on average,
see Figure \ref{fig:longitudeb}.
The single-gate errors for both the analytic and numerical gates are
less than $10^{-4}$, which makes them sufficient for quantum error correction--a
prerequisite for fault-tolerant quantum
computing \cite{aharonov2008fault, knill2005quantum, gottesman1997stabilizer}.
However, the numerical gates achieve single-gate errors
$\sim 5$ times less than those for the analytic gates,
which tracks closely with their relative improvement
on the integrated depolarization rate metric, see Appendix \ref{appendix:longitude}.
This advantage in single-gate errors corresponds to a
significant reduction in error correction resources
\cite{paetznick2014resource, suchara2013comparing}.
Furthermore, for successive gate applications,
the gate error due to depolarization is approximately linear
in the gate count, which we expect for $t \ll T_{1}$, see Figure \ref{fig:longitudec}.
The gate error reduction for large gate counts is
important for noisy, intermediate-scale quantum (NISQ)
applications. These improvements are significant for the constraints \textcolor{blue}{[improvements are significant FOR THE CONSTRAINTS?]}
we have imposed on the gates,
and do not represent a fundamental limit to the optimization
methods we have employed.

### 2021-03-22 paper/5static.tex
jens all edits plus some thomas revisions

\section{Robustness to Static Parameter Uncertainty \label{sec:static}}
We have formulated the QOC
problem as an open-loop optimization problem; equivalently,
we do not incorporate feedback from the experiment in optimization.
However, the device's parameters deviate from the parameters we use in optimization,
leading to poor experimental performance. We combat errors
of this form using robust control techniques,
making the state evolution insensitive
to parameter uncertainty. As an example,
we mitigate errors arising from the drift and finite measurement
precision of the qubit frequency, which modifies the fluxonium Hamiltonian
\eqref{eq:hamiltonian} by $f_{q} \rightarrow f_{q} + \delta f_{q}$.
We consider three robust control techniques to accomplish this task:
a sampling method, an unscented sampling method,
and a derivative method.

The sampling method incentivizes the optimizer
to ensure multiple copies of a state, each of which evolves
with a distinct value of the uncertain parameter, achieve
the same target state. Variants of this technique have been proposed
in the context of QOC
\cite{allen2019robust, ball2020software, khaneja2005optimal,
  reinhold2019controlling, rembold2020introduction} and applied
experimentally \cite{carvalho2020error}.
For each initial state,
we add two sample states $\ket{\psi^{\pm}}$
to the augmented state \eqref{eq:astatecontrols}. The discrete dynamics
function \eqref{eq:dyn_con} is modified
so the sample states evolve under the fluxonium Hamiltonian \eqref{eq:hamiltonian}
with $f_{q} \rightarrow f_{q} \pm \sigma_{f_{q}}$ for a fixed
hyperparameter $\sigma_{f_{q}}$ which is the standard deviation of the qubit frequency.
We penalize the infidelities of the sample states and their target state
by adding a cost function to the objective \eqref{eq:costfun} of the form
$\sum_{k, \pm} b_{k} (1 - {\lvert \braket*{\psi^{\pm}_{k}}{\psi_{f}} \rvert}^{2})$
where $b_{k}$ is a constant we supply.
For this method, the standard orthonormal basis states are an insufficient choice
for the initial states. As an example, a $Z/2$ gate achieved by idling
at the flux frustration point $a_{k} = 0 \ \forall \ k$
will be robust to qubit frequency detunings for the initial states $\ket{0}$
or $\ket{1}$ because the infidelity metric is insensitive to global phases,
but this gate will not be robust for any other initial states.
Therefore, we choose four initial states
so that their outer products
span the operators on the Hilbert space
$\{\ket{0}, \ket{1}, (\ket{0} + i\ket{1}) / \sqrt{2},
(\ket{0} - \ket{1}) / \sqrt{2}\}$ \cite{chow2009randomized},
which we refer to as the operator basis.

@@

The derivative method penalizes the sensitivity of the state
to the uncertain parameter, which is encoded in the $l$\textsuperscript{th}-order
state derivative $\ket*{\partial_{f_{q}}^{l} \psi} \equiv \partial_{f_{q}}^{l} \ket*{\psi}$.
In the $m$\textsuperscript{th}-order
derivative method, we append all state derivatives of order $1, \dots, m$
to the augmented state vector \eqref{eq:astatecontrols}
for each initial state.
We obtain the state derivatives at each time step by performing forward-mode
differentiation on the TDSE \eqref{eq:tdse}.
For example, the dynamics for the $1$\textsuperscript{st}-order derivative method are:
\begin{align}
  i \hbar \frac{d}{dt} \ket{\psi} &= H \ket{\psi},\\
  i \hbar \frac{d}{dt} \ket{\partial_{f_{q}}\psi} &=
  H \ket{\partial_{f_{q}} \psi} +
  (\partial_{f_{q}} H) \ket{\psi}.
  \label{eq:d1dyn}
\end{align}
We integrate the coupled ODEs with exponential
integrators in the discrete dynamics function \eqref{eq:dyn_con},
see Appendix \ref{appendix:derivative}.
While the state $\ket*{\psi}$ has unit norm,
the state derivatives $\ket*{\partial^{l}_{f_{q}} \psi}$ need not, as is evident
from the non-unitary dynamics \eqref{eq:d1dyn}.
We penalize the norms of the state derivatives
in the objective \eqref{eq:costfun} by setting the corresponding elements
of the target augmented state to zero. Intuitively, this corresponds to penalizing
the sensitivity of each state element to the uncertain parameter. As was the case for
the unscented sampling
method, we do not observe a performance increase when using more than one initial state
for the derivative method.
For runtimes of the three robust control techniques,
consult Appendix \ref{appendix:time}.

>>>>

\section{Robustness to Static Parameter Uncertainty \label{sec:static}}
We have formulated the QOC
problem as an \emph{open-loop} optimization problem, i.e.,
we do not incorporate feedback from the experiment into the optimization.
However, the precise device parameters will realistically deviate from the parameters we use in optimization,
leading to poor experimental performance. We combat errors
of this form using robust control techniques,
making the state evolution insensitive
to parameter uncertainty. As an example,
we mitigate errors arising from the drift and finite measurement
precision of the qubit frequency, which modifies the fluxonium Hamiltonian
\eqref{eq:hamiltonian} by $f_{q} \rightarrow f_{q} + \delta f_{q}$.
We consider three robust control techniques to accomplish this task:
a sampling method, an unscented sampling method,
and a derivative method.

\textcolor{blue}{[Consider structuring the following into subsections to clearly separate the three methods.]}
The sampling method incentivizes the optimizer
to ensure that multiple copies of a state, each evolving
with a distinct value of the uncertain parameter, achieve
the same target state. Variants of this technique have been proposed
in the context of QOC
\cite{allen2019robust, ball2020software, khaneja2005optimal,
  reinhold2019controlling, rembold2020introduction} and applied
experimentally \cite{carvalho2020error}.
For each initial state,
we add two sample states $\ket{\psi^{\pm}}$
to the augmented state \eqref{eq:astatecontrols}. The discrete dynamics
function \eqref{eq:dyn_con} is modified
so the sample states evolve under the fluxonium Hamiltonian \eqref{eq:hamiltonian}
with $f_{q} \rightarrow f_{q} \pm \sigma_{f_{q}}$ for a fixed
 standard deviation $\sigma_{f_{q}}$ of the qubit frequency, acting as a hyperparameter.
We penalize the infidelities of the sample states with respect to the target state
by adding a cost function to the objective \eqref{eq:costfun} of the form
$\sum_{k, \pm} b_{k} (1 - {\lvert \braket*{\psi^{\pm}_{k}}{\psi_{f}} \rvert}^{2})$
where $b_{k}$ is a constant we supply.
For this method, the standard orthonormal basis states are an insufficient choice
for the initial states. As an example, a $Z/2$ gate achieved by idling
at the flux frustration point ($a_{k} = 0 \ \forall \ k$)
will be robust to qubit frequency detunings for the initial states $\ket{0}$
or $\ket{1}$ because the infidelity metric is insensitive to global phases,
but this gate will not be robust for any other initial states.
Therefore, we choose the four initial states $\{\ket{0}, \ket{1}, (\ket{0} + i\ket{1}) / \sqrt{2},
(\ket{0} - \ket{1}) / \sqrt{2}\}$ \cite{chow2009randomized},
whose outer products span the operators on the Hilbert space, and  we refer to them as the operator basis. \textcolor{blue}{[I cannot find the term ``operator basis" in the rest of the paper. If not used, then don't introduce terminology.]}

@@

The derivative method penalizes the sensitivity of the state
to the uncertain parameter, which is encoded in the $l$\textsuperscript{th}-order
state derivative $\ket*{\partial_{f_{q}}^{l} \psi} \equiv \partial_{f_{q}}^{l} \ket*{\psi}$.
In the $m$\textsuperscript{th}-order
derivative method, we append all state derivatives of order $1, \dots, m$
to the augmented state vector \eqref{eq:astatecontrols}
for each initial state.
We obtain the state derivatives at each time step by performing forward-mode
differentiation on the TDSE \eqref{eq:tdse}.
For example, the dynamics for the $1$\textsuperscript{st}-order derivative method are:
\begin{align}
  i \hbar \frac{d}{dt} \ket{\psi} &= H \ket{\psi},\\
  i \hbar \frac{d}{dt} \ket{\partial_{f_{q}}\psi} &=
  H \ket{\partial_{f_{q}} \psi} +
  (\partial_{f_{q}} H) \ket{\psi}.
  \label{eq:d1dyn}
\end{align}
We integrate the coupled ODEs with exponential
integrators in the discrete dynamics function \eqref{eq:dyn_con} \textcolor{blue}{[integrators IN A function? sounds like mix-up between the function vs. how it is implemented numerically]},
see Appendix \ref{appendix:derivative}.
While the state $\ket*{\psi}$ has unit norm,
the state derivatives $\ket*{\partial^{l}_{f_{q}} \psi}$ need not, as is evident
from the non-unitary dynamics \eqref{eq:d1dyn}.
We penalize the norms of the state derivatives
in the objective \eqref{eq:costfun} by setting the corresponding elements
of the target augmented state to zero. Intuitively, this corresponds to penalizing
the sensitivity of each state element to the uncertain parameter. As was the case for
the unscented sampling
method, we do not observe a performance increase when using more than one initial state
for the derivative method.
For runtimes of the three robust control techniques \textcolor{blue}{[a ``technique" is abstract/general; a technique does not have a concrete runtime; do you mean computational complexity?]},
consult Appendix \ref{appendix:time}.
